Soundness Proof of Dynamic Shortcut

### 2020-11-19 paper/pldi21/report/def.tex
Modified definition

Abstract interpretation is often defined with \textit{analysis sensitivity} to
increase the precision of static analysis.  A sensitive abstract domain
$\sabsdom: \viewset \rightarrow \absdom$ is defined with a \textit{view
abstraction} $\viewmap: \viewset \rightarrow \dom$ that provides multiple points
of views for reachable states during static analysis.  It maps a finite number
of views $\viewset$ to sets of states $\dom$. Each view $\view \in \viewset$
represents a set of states $\viewmap(\view)$.
A \textit{sensitive state abstraction} $\dom
\galois{\alpha_\viewmap}{\gamma_\viewmap} \sabsdom$ is a Galois connection between
the concrete domain $\dom$ and the sensitive abstract domain $\sabsdom$ with the
following concretization function:
\[
  \gamma_\viewmap(\sabselem) = \{ \st \in \stset \mid \forall \view \in \viewset.
  \; \st \in \viewmap(\view) \Rightarrow \st \in \gamma \circ \sabselem(\view) \}
\]

@@

@@

We define a \textit{combined domain} $\combdom = \sabsdom \times
\powerset{\symbstset}$ as a pair of a sensitive abstract domain and a set of
sealed symbolic states.   The join operator of combined states is defined in an
element-wise manner:
\[
  \forall (\sabselem, S), ({\sabselem}', S') \in \combdom. \;
  (\sabselem, S) \join ({\sabselem}', S') = (\sabselem \join {\sabselem}', S \cup S')
\]

To freely convert a pair of view and an abstract state to its
corresponding symbolic state and vice versa, we define two domain converters
$\symbstset \galois{\saconverter}{\asconverter} (\viewset \times \absdom)$.  In
the converter $\asconverter$ from abstract states to symbolic states, if an
abstract value $\absval$ represents a singleton value, the function transforms
the abstract value to the corresponding concrete value.  Otherwise, it converts
the abstract value as a symbolic value in a result symbolic state.  Two
converters convert given elements without loss of information:
\[
  \saconverter \circ \asconverter = \asconverter \circ \saconverter = \identity
\]
where $\identity$ denotes the identity function.

An analysis element $\aelem \in \aelemset = (\viewset \times \absdom) \uplus
\symbstset$ is either a pair of a view and an abstract state or a sealed
symbolic state.  We define a $\atriage$ function using two converters between
analysis elements as follows:
\begin{definition}[$\atriage$]
  The triage function $\atriage: \aelemset \rightarrow \aelemset$ for analysis
  elements is defined as
  \[
    \atriage(\aelem) = \left\{
      \begin{array}{ll}
        \asconverter(\aelem)
        & \text{if} \; \aelem = (\view, \abselem) \wedge \checker(\aelem)\\
        \saconverter(\aelem)
        & \text{if} \; \aelem = \symbst \wedge \symbst \symbtrans \symbst'\\
        \aelem
        & \text{Otherwise}
      \end{array}
    \right.
  \]
  where the function $\checker$ decides whether to perform sealed symbolic
  execution instead of abstract interpretation.
\end{definition}
Using the $\atriage$ function for analysis elements, we defined the $\triage$
function for combined states:
\begin{definition}[$\triage$]
  The triage function $\triage: \combdom \rightarrow \combdom$ for combined
  states is defined as
  \[
    \triage((\sabselem, S)) = \left(
      \lambda \view. \;  \bigjoin \{ \abselem \mid (\view, \abselem) \in E \},
      E \cap \symbstset
    \right)
  \]
  where
  \[
    E = \dot{\atriage}(\{ (\view, \abselem) \mid \view \in \viewset \wedge
    \sabselem(\view) = \abselem \} \cup S)
  \]
\end{definition}
The dot notation $\dot{f}$ denotes the element-wise extended function of a
function $f$.

We could configure when the sealed symbolic execution or the abstract
interpretation is performed based on the definition of the $\checker$ function.
For example, we could define the $\checker$ that passes only function entry,
call, and exit points to perform conversions in a function level.  For the
soundness and the termination of the the abstract interpretation with dynamic
shortcuts, it should satisfy the following condition:
\begin{theorem}\label{theorem:shortcut}
  The abstract interpretation with dynamic shortcut is sound and terminates in a
  finite time if its abstract semantics is sound and $\checker$ satisfies the
  following condition:
  \[
    \checker((\view, \abselem)) \Rightarrow \asconverter((\view, \abselem))
    \symbtrans^k \excst \wedge 1 < k \leq N
  \]
  where $N$ is a pre-defined maximum length of the sealed symbolic execution.
\end{theorem}

>>>>

Abstract interpretation is often defined with \textit{analysis sensitivity} to
increase the precision of static analysis.  A sensitive abstract domain
$\sabsdom: \viewset \rightarrow \absdom$ is defined with a \textit{view
abstraction} $\viewmap: \viewset \rightarrow \dom$ that provides multiple points
of views for reachable states during static analysis.  It maps a finite number
of views $\viewset$ to sets of states $\dom$. Each view $\view \in \viewset$
represents a set of states $\viewmap(\view)$.
A \textit{sensitive state abstraction} $\dom
\galois{\alpha_\viewmap}{\sgamma} \sabsdom$ is a Galois connection between
the concrete domain $\dom$ and the sensitive abstract domain $\sabsdom$ with the
following concretization function:
\[
  \sgamma(\sabselem) = \{ \st \in \stset \mid \forall \view \in \viewset.
  \; \st \in \viewmap(\view) \Rightarrow \st \in \gamma \circ \sabselem(\view) \}
\]

@@

We also define an \textit{abstract instantiation map} $\absimap \in \absimapset:
\symbset \rightarrow \absvalset$ with abstract values $\absvalset$.  Its
concretization function $\imapgamma: \absimapset \rightarrow
\powerset{\imapset}$ is defined with the concretization function $\valgamma:
\absvalset \rightarrow \powerset{\valset}$ for values as follows:
\[
  \imapgamma(\absimap) = \{
    \imap \mid \forall \symb \in \symbset. \;
    \imap(\symb) \in \gamma \circ \absimap(\symb)
  \}
\]
We define the instantiation of a given sealed symbolic state $\symbst \in
\symbstset$ with an abstract instantiation map $\absimap \in \absimapset$ as
follows:
\[
  \instant{\symbst}{\absimap} = \{ \instant{\symbst}{\imap} \mid \imap \in
  \imapgamma(\absimap) \}
\]

\begin{definition}[Sealed Symbolic Domain]
  We define \textit{sealed symbolic domain} $\symbdom: \powerset{\absimapset
  \times \symbstset}$ with the following concretization function $\symbgamma:
  \symbdom \rightarrow \dom$ and sealed symbolic one-step execution $\symbstep:
  \symbdom \rightarrow \symbdom$ as follows:
  \begin{itemize}
    \item \(
      \symbgamma(\symbelem) = 
      \bigcup \{ \instant{\symbst}{\absimap} \mid (\absimap, \symbst) \in
      \symbelem\}
    \)
    \item \(
      \symbstep(\symbelem) = \{ (\absimap, \symbst') \mid (\absimap, \symbst)
      \in \symbelem \wedge \symbst \symbtrans \symbst' \}
    \)
  \end{itemize}
\end{definition}

@@

\begin{definition}[Combined Domain]
  We define a \textit{combined domain} as $\combdom = \sabsdom \times \symbdom$
  with the following concretization function and join operator:
  \begin{itemize}
    \item \(
      \combgamma((\sabselem, \symbelem)) = \sgamma(\sabselem) \cup
      \symbgamma(\symbelem)
    \)
    \item \(
      (\sabselem, \symbelem) \join ({\sabselem}', \symbelem') = (\sabselem \join
      {\sabselem}', \symbelem \cup \symbelem')
    \)
  \end{itemize}
\end{definition}

An analysis element $\aelem \in \aelemset = (\viewset \times \absdom) \uplus
(\absimapset \times \symbstset)$ is either a pair of a view and an abstract
state or a sealed symbolic state.  To freely convert a pair of a view and an
abstract state to its corresponding pair of an abstract instantiation map and a
sealed symbolic state, we define two domain converters: \begin{itemize} \item
$\asconverter: (\viewset \times \absdom) \hookrightarrow (\absimapset \times
\symbstset)$ \item $\saconverter: (\viewset \times \absdom) \leftarrow
(\absimapset \times \symbstset)$ \end{itemize} The converter $\asconverter$ is a
\textit{partial function} and it is applicable when it is possible to represent
a given abstract state in a specific view as a pair of abstract instantiation
map and a sealed symbolic state.  Two converters should convert given elements
without loss of information:
\[
  \forall \view \in \viewset, \abselem \in \absdom.
\]
\[
  \asconverter((\view, \abselem)) = (\absimap, \symbst)
  \Rightarrow \left\{
  \begin{array}{l}
    (\view, \abselem) = \saconverter((\absimap, \symbst))\\\
    \viewmap(\view) \cap \gamma(\abselem) = \instant{\symbst}{\absimap}\\
  \end{array}
  \right.
\]
where $\identity$ denotes the identity function.

We define a $\atriage$ function using two converters between analysis elements
as follows:
\begin{definition}[$\atriage$]
  The triage function $\atriage: \aelemset \rightarrow \aelemset$ for analysis
  elements is defined as
  \[
    \atriage(\aelem) = \left\{
      \begin{array}{ll}
        \asconverter(\aelem)
        & \text{if} \; \aelem = (\view, \abselem) \wedge \aelem \in
        \Dom(\asconverter)\\
        \saconverter(\aelem)
        & \text{if} \; \aelem = (\absimap, \symbst) \wedge \symbst \symbtrans \symbst'\\
        \aelem
        & \text{Otherwise}
      \end{array}
    \right.
  \]
\end{definition}
Using the $\atriage$ function for analysis elements, we defined the $\triage$
function for combined states:
\begin{definition}[$\triage$]
  The triage function $\triage: \combdom \rightarrow \combdom$ for combined
  states is defined as
  \[
    \triage((\sabselem, \symbelem)) = \left(
      \lambda \view. \bigjoin \{ \abselem \!\mid\! (\view, \abselem) \in E \},
      E \cap (\absimapset \times \symbstset)
    \right)
  \]
  where
  \[
    E = \dot{\atriage}(\{ (\view, \abselem) \mid \view \in \viewset \wedge
    \sabselem(\view) = \abselem \} \cup \symbelem)
  \]
\end{definition}
The dot notation $\dot{f}$ denotes the element-wise extended function of a
function $f$.

\begin{definition}[Combined One-Step Execution]
  We define the combined one-step execution $\combstep: \combdom \rightarrow
  \combdom$ is define as follows:
  \[
    \combstep(\combelem) = (
      \sabsstep(\sabselem),
      \symbstep(\symbelem)
    )
  \]
  where $(\sabselem, \symbelem) = \triage(\combelem)$.
\end{definition}

We could configure when the sealed symbolic execution or the abstract
interpretation is performed based on the definition of the converter
$\asconverter$.  For example, we could define the $\asconverter$ for analysis
elements related to only function entry, call, and exit points to perform
conversions in a function level.  For the soundness and the termination of the
the abstract interpretation with dynamic shortcuts, it should satisfy the
following condition:
\begin{theorem}\label{theorem:shortcut}
  The abstract interpretation with dynamic shortcut is sound and terminates in a
  finite time if its abstract semantics is sound and $\saconverter$ satisfies the
  following condition:
  \[
    \asconverter((\view, \abselem)) = (\absimap, \symbst) \Rightarrow
    \symbst \symbtrans^k \excst \wedge 1 < k \leq N
  \]
  where $N$ is a pre-defined maximum length of the sealed symbolic execution.
\end{theorem}
